### 2015-11-12 ch02-qlogms/excitation-design.tex
Beunen in excitation design

\section{\TODO{Robust vs optimal}}
\section{\TODO{Robust QLog MS}}
\subsection{\TODO{Constraints: time and power}}
\subsection{\TODO{Informativity of experiment}}
\subsection{\TODO{Normalized design w.r.t. damping}}
\subsection{\TODO{Grid misalignment effects}}
\subsection{\TODO{qLog design rules}}

>>>>

\TODO{reshape figures}
\TODO{rename labels}

\section{Introduction} \label{sec:intro}
  It often occurs that we wish to measure the \gls{FRF} of a \gls{LTI} system in a frequency band that covers several decades~\citep{Bragos2001,Sanchez2011,Niedostatkiewicz2009,VanGheem2004}.
  Of course, we also want to attain a specified uncertainty level over the full frequency span.

  Any LTI system of finite order can be expanded into a  finite sum of systems of first and second order by means of a partial fraction expansion\citep{Oppenheim1983}.
  Therefore the performance of the proposed method is assessed on a prototype system of second order.
  This does not limit the generality of the approach as we can add several of such systems to construct a general LTI system.

  To ensure that any second order system with its resonance frequency in the specified frequency range is measured with an equal accuracy, the system output should contain a comparable power level irrespective of its resonance frequency.
  For a system with a sufficiently low damping, one can intuitively understand that most of the information is contained within the $3 \unit{dB}$ bandwidth of the system.
  Remember that an LTI system is naturally described in a logarithmic frequency axis (the slopes of such a system are specified in $\unit{dB}/\text{decade}$).
  Hence, an excitation with a flat power spectrum and a linearly spaced frequency grid will over-excite the high end of the band at the cost of the low frequencies.
  To restore equal excitation signal levels over the complete range of frequencies, we will need to distribute the power in $1/\omega$ over the frequency band.
  In a noise excitation context, this results in pink noise.
  In a periodic setting, this results in a multisine whose frequency components are spaced equidistantly on a logarithmic axis.

  The main problem is that network and dynamic signal analyzers typically
  rely on an equidistant frequency grid that comes with the Discrete Fourier Transform (DFT).
  A logarithmically spaced signal, on the other hand, does not match well to an equidistant frequency grid.
  More specifically, the lack of frequency resolution at the lowest frequencies hampers the quality of the power distribution at the low edge of the frequency band.
  To circumvent this disadvantage, we will construct quasi-logarithmically spaced signals.
  Their frequency spacing is approximately logarithmic but nevertheless conforms to the DFT's linear frequency grid.
  This problem was already discussed in \citet{Geerardyn2012IMTC}.

  We will also present a simple method to determine a suitable density of a logarithmically spaced signal for a predefined minimal damping of the system to be tested.

  In the first section of this paper, we start from a generic transfer function model and discuss a decomposition thereof in second order subsystems and their properties.
  We provide a comparison between several possible implementations of the frequency grids.
  We discuss their applicability to multisine signals.
  We also introduce a compensation method for the problems of the quasi-logarithmic grid and propose a novel $1/f$ `pink' quasi-logarithmic multisine.
  The second section describes the selection of a suitable frequency ratio $\alpha$ for a logarithmically spaced grid such that constant quality (signal to noise ratio (SNR)) is obtained.

  In the third section, the presented techniques are illustrated on a simulation example.
  Finally, we verify our results experimentally on the measurements of a band-pass filter.

\section{Methodology} \label{sec:method}

  \subsection{The system under consideration}
  We consider the transfer function models of generic \gls{LTI} systems of arbitrary but finite orders $n_a\times n_b$:
  \begin{align}
    G\left( \Omega, \theta \right) = \frac{B\left( \Omega,\theta \right)}
                                          {A\left( \Omega,\theta \right)}
                                   = \frac{\sum\limits_{i=0}^{n_b-1} b_i \Omega^i}
                                          {\sum\limits_{i=0}^{n_a-1} a_i \Omega^i}
    \text{.} \label{eq:model}
  \end{align}

  The model is parametrized in the vector $\theta = \left[b_0 \,\cdots\, b_{n_b-1} \; a_0 \,\cdots\, a_{n_a-1} \right]^{\T}$ and  is evaluated at different frequencies $\Omega$ in the complex plane.
  For a continuous time model $\Omega = s$, while for the discrete time case $\Omega = z$.

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion.

  To study the behavior of one sub-system, we inspect a generic second order system $G\left( s \right)$.
  In continuous time such a system can be described by a transfer function of the form
  \begin{equation}
    G\left(s \right) = \frac{\dc}
                            {             \frac{s^2}
                                               {\wn^2} 
                             + 2 \damping \frac{s}
                                               {\wn}
                             + 1
                            }
  \text{.}
  \label{eq:secondOrderSystemCT}
  \end{equation}
\TODO{This is actually wrong, rather, it should be $\frac{as-b}{cs^2 + ds + 1}$. However, the difference in information is rather small around the actual resonance. See script in NnF folder.}
\TODO{write out in partial fraction expansion, use script to show that in practice, only a few parameters matter. This will be well-approximated by considering only the second order system of the paper.}

  The DC gain $\dc$, the (relative) damping $\damping$ and the natural pulsation $\wn$ are a more intuitive parametrization for this system than the generic polynomial coefficients $a_i$ and $b_i$ used in \eqref{eq:model}.
  The corresponding resonance pulsation is $\wR \isdef \wn \sqrt{1-\damping^2}$.
  The $3\unit{dB}$ bandwidth around the resonance is given by $\wdB = 2\damping\wn$ for systems that have a sufficiently low damping.

  When $\damping < \frac{1}{\sqrt{2}}$, such a second order system exhibits a peak in
  the amplitude of the transfer function \citep{Oppenheim1983}:
  \begin{align}
   \left| G\left(\wM \right) \right| & \isdef
          \max_{\omega} \left| G\left( \omega \right)  \right|
        = \frac{\dc}
               {2 \damping \sqrt{1-\damping^2}}
               \label{eq:peakAmplitude}\\
   \wM &
        = \sqrt{1-2\damping^2} \wn
        \label{eq:peakFrequency}
    \text{.}
  \end{align}

\subsection{System and noise model}
  In this paper, an output error (OE) system setup is used.
  Denote
    $U  \left(\omega\right)$ as the excitation signal,
    $Y_0\left(\omega\right)$ the exact, unperturbed system output,
    $E  \left(\omega\right)$ for the output noise source,
    $Y  \left(\omega\right) = Y_0\left( \omega \right) + E\left( \omega \right)$ depicts the measured output signal.

  The output noise is assumed to be zero-mean and white with a
  standard deviation of $\sigma_e$ in the time domain or $\sigma_E$ in the frequency domain.

  \TODO{put block diagram here}

\subsection{Multisine Excitations}
  To excite the system under test we consider a generalization of multisine excitations \citep{Pintelon2012}:
  \begin{definition}[Generalized multisine] \label{def:generalized-MS}
  A generalized multisine $u(t)$ is a signal consisting of $F$ sine waves with a different frequencies $f_k$, amplitudes $A_k$ and phase shifts $\phi_k$:
  \begin{equation}
    u \left( t\right) = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(2\pi f_k t + \phi_k \right)
  \text{.}
  \label{eq:MultisineCT}
  \end{equation}
  For simplicity, one can stipulate that the frequency grid is sorted, i.e. $f_1 < f_2 < \ldots < f_F$.
  \end{definition}

  We limit ourselves, without loss of generality, to random phase multisines.
  Hence, $\phi_k$ is the outcome of a uniform random process over $\left[0,2\pi\right[$.
  This leaves two sets of free parameters to create a signal that suits our needs:
    the amplitude of the spectral lines $A_k$ and
    the frequency grid $\left\{f_1, \ldots, f_F \right\}$.
  Note that the $k^{\text{th}}$ term in \eqref{eq:MultisineCT} is has a period $T_k \isdef f_k^{-1}$.

  Consequently, the complete multisine has a period $T_{u(t)}$ that is the common multiple of the period of all its components:
  \begin{equation}
    T_{u(t)}  = \lcm \set{T_1, \ldots, T_F}
      = \lcm \set{f_1^{-1}, \ldots, f_F^{-1}}
    \text{.}
  \end{equation}
  \begin{definition}[Least common (integer) multiple] \label{def:lcm-real}
  The least common multiple of the set of values $T_k \in \RealNumbers$ for $k \in \set{1, \ldots, F}$ is  
  \begin{equation}
    \lcm \set{T_1, \ldots, T_F} \isdef
      \min \set{T \given{\forall k \in \set{1,\ldots,F}, \exists n_k \in \NaturalNumbers: n_k T_k = T }}
   \end{equation}
   if such a minimum exists.
   \end{definition}
  \begin{example}
   Consider a generalized multisine where $f_1 = 1 \unit{Hz}$ and $f_2 = \tfrac{2}{3} \unit{Hz}$.
   The corresponding periods are $T_1 = f_1^{-1} = 1 \unit{s}$ and $T_2 = f_2^{-1} = 1.5 \unit{s}$.
   The overall period of the multisine is $T = n_1 T_1 = n_2 T_2 = 3 \unit{s}$, i.e. $n_1 = 3$ and $n_2=2$.
  \end{example}

  \begin{example}
   Consider the generalized multisine of the previous example to which a spectral component is added at $f_3 = \pi \unit{Hz}$, or equivalently one with a period of $T_3 = \pi^{-1} \unit{s}$.
   Then it is not possible to determine integer values $n_1$, $n_2$, $n_3$ such that $n_1 T_1 = n_2 T_2 = n_3 T_3$.
   Consequently, such a `multisine' is not periodic.
  \end{example}

  This last example illustrates that a least common multiple of real numbers does not always exist.
  Consequently, the generalized multisine of \defref{def:generalized-MS} is not guaranteed to be periodic.

  \TODO{put MATLAB code for each type of signal}

\subsection{Equidistant (Linear) Grid Multisine}
  In many applications, an equidistant frequency grid is used for $u\left( t\right) $ because this fits best with the classical DFT analysis \citep{OppenheimDT,Mandal2007}. It allows to analyze periodic signals on an equidistant frequency grid very efficiently.

  An equidistant frequency grid $\set{f_1, \ldots, f_F}$ with spacing $\Delta f$
  consisting of $F$ lines conforms to the following relation:
  \begin{equation}
    f_k = k \Delta f + k_0 \Delta f
    \emspace \forall k \in \set{1,\ldots,F},
    \emspace k_0 \in \NN
    \text{.}
  \label{eq:linGrid}
  \end{equation}

  Using such linear grid, we can easily construct a representation of a multisine
  in discrete time that is suited for use with the DFT by sampling at a given sampling rate $f_s$:
  \begin{equation}
     u \left[ n\right] = \frac{1}{\sqrt{F}}
     \sum_{k=1}^{F} 
       A_k 
       \sin 
         \left(\frac{2\pi n \left( k + k_0 \right) \Delta f}{f_s} + \phi_k \right)
    \label{eq:MultiSineDT}
  \end{equation}

  Such an equidistant grid is a very common choice.
  Note that it has more frequency lines per decade (or octave) at higher frequencies.
  Therefore, the excitation power per octave is larger at higher frequencies
  when a constant amplitude spectrum ($A_k = A_{\mathrm{in}}$) is used.
  For a constant amplitude spectrum, this means the dynamics that are present at lower frequencies are plagued by a higher uncertainty than the ones located at higher frequencies.

  Therefore a signal with an equidistant frequency grid is undesirable for
  our situation as it wastes signal power at high frequencies at the cost of an increased variance in the lower side of the band.

\subsection{(Ideal) Logarithmic Grid Multisine}
  For a logarithmic frequency grid $\set{f_1, \ldots, f_F}$ of $F$ lines,
  the following relation holds between excited frequencies:
  \begin{equation}
    f_k = \alpha \cdot f_{k-1}
    \emspace \forall k \in \set{1,\ldots,F}
  \label{eq:logGrid}
  \end{equation}
  for a given lowest frequency $f_0$ and frequency ratio $\alpha > 1$.

  When designing an excitation signal, it is common only to specify a (large) frequency band
  by its boundary frequencies $f_{\min}$ and $f_{\max}$.
  For a given $\alpha$ it is easy to determine the number of excited lines
  within the frequency range of interest:
    \begin{equation}
      F = \floor{
                  \frac{\log{f_{\max}} - \log{f_{\min}}}
                       {\log{\alpha}}
                }
    \label{eq:logms:F}
    \end{equation}
  where $\floor{\bullet}$ denotes rounding towards zero.
  In the Section \ref{sec:optimAlpha}, the choice of a suitable $\alpha$ is elaborated.

  \TODO{link naar log meas mode DSA (e.g. HP3562A) leggen. Bij HP 3562A, is $F/\mathrm{decade} = 80$, oftewel $\alpha = \sqrt[80]{10} \approx 1.029$ vast ingesteld. Dit beperkt bruikbaarheid van zo'n toestel.}

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B}(f) \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}

   Obviously, the same argument can be held for any set of frequency bands with constant relative bandwidth (e.g. octave $\kappa = 2$).
   I.e. the same number of frequency components is present per decade, such that the power in each decade is identical when a constant amplitude spectrum ($A_k = A_{\mathrm{in}}$) is used.

  To return to the FRF, in Section~\ref{sec:optimAlpha} we will show that for a dense logarithmic frequency grid, any second order system with damping $\damping$ and a resonance within the bulk of the frequency grid, will receive the same number of frequency lines in its $3 \unit{dB}$ bandwidth regardless of the actual value of the resonance frequency.
  As most of the information of such systems is obtained from the measurements at frequencies inside the $3 \unit{dB}$
  bandwidth, one expects an equal variance for each system, as will be shown in Section~\ref{sec:bestFrequencyResolution}.

  For that reason, a log multisine is a good candidate for our purposes.
  However, it is more involved both to generate and to analyze such a signal, as the excited frequencies don't lie on a commensurate frequency grid.
  Especially for periodic measurements, this is a major downside since the period of such signal does not always exist (i.e. this leads to infinite measurement times).
  It is therefore not practical to use a perfectly logarithmically spaced frequency grid as the \gls{DFT} or \gls{CZT}~\citep{Rabiner2004} explicitly rely on a commensurate frequency grid.

\subsection{Quasi-Logarithmic Grid Multisine}
  Since most measurement equipment and analysis methods rely on signals with an equidistant frequency grid, we create a frequency grid
  $\set{f_1, f_2, f_3,\ldots, f_F}$ for which \eqref{eq:logGrid} is approximately  valid but with the strict constraint that the frequencies must be a subset  of an equidistant frequency grid. 
  This yields the following relation for  subsequent frequency lines:
  \begin{equation}
    f_k \approx \alpha \cdot f_{k-1}
    \emspace \forall k \in \set{1,\ldots,F}
  \label{eq:qlogGrid}
  \end{equation}
  under the constraint that
  \begin{equation}
    f_k = N \cdot \Delta f,\; N \in \NN
    \text{.}
    \label{eq:constraintLinGrid}
  \end{equation}
  Such a grid is called a quasi-logarithmic (quasi-log) grid~\citep{Pintelon2012}.

  \begin{remark}
  When $\alpha \not\in \NaturalNumbers$, it is possible that at the low end of the frequency band, $\Delta f > \left(\alpha -1 \right) f_{k-1}$ for some integer values of $k$.
  In such cases, the commensurate frequency grid is not dense enough to realize a frequency ratio $\alpha$.
  Consequently, the power spectrum at these low frequencies will be less than the one of an ideal logarithmic frequency grid.
  \end{remark}

  \begin{definition}
  The effective frequency ratio $\alpha_k$ of a frequency grid is
  \begin{equation}
    \alpha_k \isdef  \frac{f_k}{f_k-1}
    \text{.}
  \end{equation}
  \end{definition}

  \begin{remark}
  For the implementation of a quasi-logarithmic frequency grid,  the approximate equality  \label{eq:qlogGrid} does not define the grid unambiguously.
  The quasi-logarithmic grids constructed in this chapter are constructed using the relationship
  \begin{equation}
    f_k = \alpha_k f_{k-1} 
    \text{ with } 
      \alpha_k \isdef 
          \round{   \frac{\alpha f_{k-1}}{\Delta f}}    
           \frac{\Delta f}{f_{k-1}}
      \approx \alpha 
      \emspace 
          \forall k \in \set{1,\ldots,F}
     \text{.}
     \label{eq:excitation:qlog:round}
  \end{equation}
  In the equation above, $\round{\placeholder}$ denotes rounding towards the nearest integer as in the \MATLAB function \code{round}.
  This satisfies the approximate definition.
  \end{remark}

  \begin{remark}
  The use of different rounding strategies in \eqref{eq:excitation:qlog:round}, allows for slightly different guarantees on the effective $\alpha_k$ attained by the grid.
  By denoting $\alpha_k^{\placeholder}$ as the effective frequency ratio for a rounding strategy $\placeholder$  in that equation, one can trivially see that the following relationships hold:
  \begin{equation}
  \alpha_k^{\floor{}} \leq \alpha \leq  \alpha_k^{\ceil{}}
  \qquad \text{and} \qquad
  \alpha_k^{\floor{}} \leq \alpha_k^{\round{}} \leq  \alpha_k^{\ceil{}}
  \qquad
  \forall k \in \set{1,\ldots,F}
  \text{.}
  \end{equation}
  \end{remark}

  As this signal approximates a logarithmically spaced signal,  we expect that it will exhibit comparable properties.
  On the other hand, by imposing \eqref{eq:constraintLinGrid}, it remains
  possible to use signal techniques which rely on a commensurate frequency grid.

\subsection{Compensated Quasi-Logarithmic Grid Multisine}
  The power density of the quasi-log multisine at the low frequencies is reduced in comparison to that of the logarithmic grid multisine.
  This is due to the limited available frequency resolution.
  To compensate for this loss in power spectral density (PSD), we can increase the amplitude spectrum $A_k$ at these lines.
  Note that we are not able to restore the frequency resolution, only the power in a certain band is made equal to the power in the logarithmic multisine in the same band.

  First of all, we need to determine the factor $A_k$ needed to correct the power spectrum to match that of a log grid multisine.
  To this end, one first constructs a log frequency grid $\LOGGRID = \set{\f_1, \ldots, \f_{F'}}$ for a given factor $\alpha$ and the corresponding quasi-log grid $\QLOGGRID = \set{f_1, \ldots, f_F}$ for a resolution $\Delta f$.
  For each frequency $f_k$ in the quasi-log grid, we define $n\left(f_k\right)$
  as the number of frequencies in the log grid that are nearer to $f_k$
  than to any other grid line in the quasi-log grid:

  \begin{equation}
    n\left(f_k\right) = \numberOf \mathrm{nearestLines}_{f_k}
    \label{eq:nfk}
  \end{equation}
  \begin{equation}
    \mathrm{nearestLines}_{f_k} =
      \left\{ 
        \f_l \given{\begin{aligned} 
                       \f_l \in \LOGGRID, \\
                       \norm{\f_l - f_k} < \norm{\f_l - f_{k'}},\\
                       \forall k' \neq k: f_k, f_{k'} \in \QLOGGRID
                    \end{aligned}}
      \right\}
    \label{eq:nearestLines}
  \end{equation}

  When approximating a log multisine with constant amplitude spectrum $A_{\mathrm{in}}$, we choose the amplitude spectrum of the compensated quasi-log multisine as follows
  \begin{equation}
    A_k = A_{\mathrm{in}} \cdot \sqrt{n\left(f_k\right)}\emspace \forall k \in \set{1,\ldots,F}
  \text{.}
  \label{eq:compensationAk}
  \end{equation}
  This effectively concentrates the total power of the $n\left( f_k \right)$ surrounding lines in the log grid multisine at the frequency $f_k$ in the compensated quasi-log multisine.

  The power spectrum of this compensated signal approximates the one of the log grid multisine better than an uncompensated quasi-log grid.
  As the power spectrum approximates the power spectrum of the log grid multisine, the uncertainty of the frequency response function will be approximately independent of the frequency within the measured frequency band.

  \subsection{Pink Quasi-Logarithmic Multisine}
  In \citep{Rojas2007} it is suggested that band-limited $1/f$ noise is a reasonable, albeit not optimal, robust excitation signal.

  For $1/f$ noise (`pink noise'), the power spectral density $S(f)$ is proportional to $1/f$.
  We can again approximate such a signal by means of a quasi-logarithmic grid multisine.
  To do so, we construct a quasi-logarithmic frequency grid $\set{f_1,\ldots,f_k,\ldots,f_F}$ and determine the corresponding amplitude spectrum $A_k$ such that each excited line carries the same power as the corresponding frequency band does for pink noise.

  This allows to define the amplitude spectrum $A_k$ as
  \begin{equation}
    A_k = A_{\mathrm{in}}
                 \sqrt{\int_{\fMin{k}}
                     ^{\fMax{k}}
                     \limits
                     \frac{1}
                          {f}
                     \,\dd{f}}
          = A_{\mathrm{in}}
                \cdot
                \sqrt{
                \ln \frac{\;\fMax{k}\;}{\fMin{k}}
                }
  \end{equation}
  where $\left[ \fMin{k}, \fMax{k} \right] $ denote the frequency range over which PSD of the pink noise is to be approximated by the power at frequency line $f_k$.

  To reduce the complexity of the expressions for $\fMin{k}$ and $\fMax{k}$, we extend the frequency grid by one component to the left and to the right by means of the grid relation (equations \eqref{eq:qlogGrid} and \eqref{eq:constraintLinGrid}) to obtain the grid $\set{f_0, f_1, \ldots, f_F, f_{F+1}}$.

  Using this extension, the frequency range covered by $f_k$ can then be calculated as
    \begin{align}
      \fMin{k} &= \mean{f_{k-1}}{f_k}\\
      \fMax{k} &= \mean{f_{k+1}}{f_k}
    \label{eq:fRange}
    \text{.}
    \end{align}
  This range can be seen as the power spectral density counterpart of the discrete spectrum relation \eqref{eq:nearestLines}.
  Note that instead of the geometric mean, the arithmetic mean can be used in \eqref{eq:fRange} to calculate almost identical boundaries.

  \subsection{Comparison of the frequency grids}
  To gain some insight into these different grids, \figref{fig:freqGrids} depicts the amplitude spectrum for a linear, log, compensated quasi-log and pink quasi-log grid multisine that all carry the same total power.
  For this example, a frequency spacing $\Delta f = 1\unit{Hz}$ and frequency ratio $\alpha = 1.05$ are used.

  \begin{figure}[ht]
    \centering
    \setlength\figureheight{5cm}
    \setlength\figurewidth{0.8\columnwidth}
    \input{\thisDir/figs/freqGrid.tikz}
    \caption[Amplitude spectrum for linear, logarithmic, quasi-logarithmic grids.]{Comparison of the amplitude spectrum of a linear grid multisine, a logarithmic grid multisine, the compensated quasi-logarithmic multisine and the `pink' quasi-log multisine.}%
    \label{fig:freqGrids}
  \end{figure}

  \begin{table}
  \centering
   \caption{Summary of the different frequency grids.}
    \begin{tabular}{lccc} 
    Grid & Linear & Logarithmic & Quasi-logarithmic \\

    % Limits & \multicolumn{3}{c}{$\ClosedInterval{\fMin, \fMax}$}\\
    Parameters & $f_0, \Delta f$ & $f_0, \alpha$ & $f_0, \Delta f, \alpha$ \\
    $f_{k+1} \isdef$ & $f_{k} + \Delta f$ & $f_{k} \cdot \alpha$ & \ldots \\

    \end{tabular}
    \TODO{afmaken, booktabs}
  \end{table}

  \TODO{plot: $\alpha $ (evt. $\cdot A_k^2$) i.f.v. $\omega$ voor verschillende signalen.}
  \TODO{plot: $\Delta f $ i.f.v. $\omega$ voor verschillende signalen} 
  \TODO{plot: $\# \omega / \mathrm{decade}$ i.f.v. $\omega$ voor verschillende signalen}

  At the lowest frequencies the quasi-log grid and the linear grid frequencies do coincide and each line in the quasi-log grid is surrounded by many lines of the log grid.
  The amplitudes of the compensated and pink quasi-log lines show that the compensation only has effect at the lower frequencies since the uncompensated counterpart has an amplitude spectrum close to $1$.
  We can expect a very similar behavior from properly designed compensated quasi-log multisines and from the `pink' quasi-log multisine, since their amplitude spectra are almost identical.

\section{Optimization of the frequency spacing} \label{sec:optimAlpha}
  \subsection{Calculation of the model uncertainty}
  In the previous section logarithmically spaced multisines were introduced, but up to now their major parameter, the frequency spacing factor $\alpha$, has not been designed for.
  In this section we show a possible approach to compare the performance of different excitation signals with respect to the theoretically minimal attainable uncertainty of the transfer function as expressed by the Cramér-Rao Lower Bound.
  Doing so will allow to determine a suitable value for the frequency spacing $\alpha$.

\TODO{zin loopt niet goed}
  The covariance matrix $\C_{G\left(\omega_{\interest}\right)}$ of the parametric transfer function $G(\theta,\omega)$ can be shown \citep{Pintelon2012, Gallager2008, matrixcookbook} to be approximately equal to
  \begin{align}
    \C_{G\left( \omega_{\interest}\right) } \approx
      \InnerQuadraticForm[\HT]
                         {\jac{\interest}}
                         {\FisherInformation_{\theta}^{-1}}
     \label{eq:varPeak}
  \end{align}
  with 
  \begin{equation}
    \FisherInformation_{\theta} =
    2
    \real{
          \InnerQuadraticForm[\HT]
                              {\jac{\excited}}
                              {\C_{X}^{-1}}
          }
    \label{eq:FisherInfo}
  \end{equation}
  the Fisher information matrix \citep{Pintelon2012}.

  In these expressions, the Jacobian $\jac{\interest}$ matrix of the transfer function evaluated at the frequency of interest $\omega_{\interest}$ with respect to the model parameters $\theta$ is defined as
  \begin{equation}
    \jac{\interest} \isdef
         \PartialDerivative{G\left( \omega_{\interest}, \theta \right) }
                           {\theta}
    \text{.}
  \end{equation}
  Similarly, $\jac{\excited}$ is the Jacobian calculated at the excited frequency lines $\omega_{\excited} = \set{\omega_1, \ldots, \omega_F}$.

  Finally, $\C_{X}$ denotes the covariance matrix of $X\left(\omega\right) = G\left(\omega, \theta_0 \right) - \frac{Y_0\left(\omega \right) + E\left( \omega \right)  }{ U\left( \omega \right) } $, evaluated at the excited frequencies $\omega_{\excited}$.
  By assuming that the $X\left( \omega \right) $ are independently distributed over the frequency $\omega$, the variance-covariance matrix $\C_{X}$ is a diagonal matrix \citep{Pintelon2012}:
  \begin{equation}
    \C_{X} =
               \left[ \left.
               \begin{matrix}
                 \frac{\sigma^2_E\left(\omega_1 \right) }
                      {\abs{ U\left(\omega_1 \right)}^2 } & 0 & 0 \\
                 0 & \ddots & 0 \\
                 0 & 0 & \frac{\sigma^2_E\left(\omega_F \right) }
                              {\abs{U\left(\omega_F \right)}^2 }
               \end{matrix}
               \right.^{\vphantom{x}}\right]
    \text{.}
  \end{equation}

  In this paper, we select a single frequency of interest $\omega_{\interest}$
  to obtain a scalar performance measure.
  This is allowed as the matrix $\C_{X}$ is diagonal.

  \subsection{Relation between the frequency grid and the system resonance} \label{sec:relationLogGridSystem}
  To determine an appropriate frequency ratio $\alpha$ for a logarithmically spaced signal, we study the estimated FRFs of a set of prototype second-order systems excited by logarithmically spaced excitation signals.

  Define $G_{1}$ for a second-order system with resonance frequency $\wn$, damping $\damping$ and DC gain $\dc$.
  Denote $G_m$ a second-order system with the same $\damping$ and $\dc$ but a resonance frequency of $m\wn$.
  If one calculates their FRFs by evaluating \eqref{eq:secondOrderSystemCT}, one sees that $G_1(\omega)$ and $G_m(m\omega)$ yield the same numerical values for any value of the pulsation $\omega$.
  Both FRFs share the same `shape', but are `located' at different frequencies.

  Consider a logarithmic frequency grid with an infinite number of grid lines
  \begin{equation}
    \vec{\omega} = \set{\alpha^k \omega_0 \given{k \in \ZN}}
    \label{eq:freqGridLog}
    \text{.}
  \end{equation}

  Consider a second grid, defined as a multiple of this grid, $m\vec{\omega} = \set{m \cdot \alpha^k \omega_0 \given{k \in \ZN}}$; both grids share the same frequency ratio $\alpha$.

  \TODO{formulate as lemmas, show steps}
  It is easy to show that for $m = \alpha^N$ the following relation holds:
  \begin{equation}
    G_1\left( \vec{\omega} \right)
    \stackrel{(a)}{=} G_{\alpha^N} \left( \alpha^N \vec{\omega} \right)
    \stackrel{(b)}{=} G_{\alpha^N} \left( \vec{\omega} \right)
    \label{eq:gridPeriodicity}
  \end{equation}
  since
  \begin{itemize}
   \item[(a)] the numerical values of the FRFs $G_1(\vec{\omega})$ and $G_m(m\vec{\omega})$ are equal and,
   \item[(b)] the grids $\alpha^N \vec{\omega} = \set{\alpha^{k+N} \omega_0 \given{k \in \ZN}}$ and $\vec{\omega} = \set{\alpha^k \omega_0 \given{k \in \ZN}}$ coincide for integer values of $N$.
  \end{itemize}

  This shows that systems with equal $\damping$ and $\dc$ but resonance frequencies that differ by a factor $\alpha^N$, will share an identical `shape` of their response at the lines of a logarithmically spaced frequency grid with infinite span and a frequency ratio $\alpha$.

  Taking this into account allows to limit the study of second order systems to a small set of systems that should be considered in the search for a suitable grid spacing $\alpha$.
  We can limit the evaluation of the theoretical uncertainty \eqref{eq:varPeak} obtained by exciting the system with a logarithmic signal with ratio $\alpha$ to the analysis of second order systems with a resonance frequency set to
  \begin{equation}
    \wn \alpha^{N+\gamma}\qquad \text{ with } 0 \leq \gamma < 1\text{.}
    \label{eq:theoreticalBestVariationOmega}
  \end{equation}
  The value $\gamma$ is varied for a fixed value of $N \in \ZN$.
  The value $N$ will not influence the `shape` (and uncertainty) of the FRF.

  The arguments for a logarithmically spaced grid with an infinite span remain approximately valid for quasi-logarithmically spaced grids with a finite frequency support if the resonance peak of the system under test can be considered within the `bulk' of the frequency grid with finite support.
  I.e. there should be a few excited frequency lines that are logarithmically spaced, outside of the resonance peak of the actual system.
  In practice, this means that systems with their resonance peak at the lowest and the very highest edge of the frequency band may not be observed with identical quality as systems at intermediate frequencies.

  \subsection{Selection of the frequency grid spacing $\alpha$}
  When we now take a closer look at the general second order system \eqref{eq:secondOrderSystemCT}, we still are left with $\dc$ and $\damping$ as free parameters in the prototypes.
  Note that $\dc$ is a scaling factor for the FRF.
  Hence the results for different values of $\dc$ only differ by a scaling factor.
  So we only need to perform the analysis for a fixed value of $\dc$, which was chosen equal to $1$ in our test case.

  The choice of the damping is more important.
  Instead of choosing the damping directly, we choose the peak gain $\left| G\left( \wM \right) \right| \in \set{5 \unit{dB}, 10 \unit{dB}, 20 \unit{dB}, 40 \unit{dB} }$ of the transfer function \eqref{eq:peakAmplitude} to study the effect of differently damped systems.

  Since the damping influences the $3\unit{dB}$ bandwidths, it will become important when one has to determine the grid spacing $\alpha$.
  To deal with this, we introduce the normalized frequency ratio
  \begin{equation}
    \tilde{\alpha} = \frac{\alpha - 1}
                          {\damping}
                   \propto \frac{\omega_{k+1} - \omega_k}
                                {\wdB}
                           \cdot
                           \frac{\wn}{\omega_k}
    \text{.}
    \label{eq:normAlpha} 
  \end{equation}
  This normalized frequency ratio can be interpreted as a measure of the grid resolution, relative to the $3\unit{dB}$ bandwidth of the second order system.
  The grid resolution in a (quasi-) log grid is $\left( \alpha - 1\right)\omega_k$ while the system bandwidth is approximately $2\damping\wn$.
  Around the resonance, $\omega_k \approx \wn$, such that the last fraction in \eqref{eq:normAlpha} can safely be neglected.

  To determine a suitable value of the frequency spacing ratio $\alpha$ for the logarithmic and (compensated) quasi-logarithmic grid, we construct a logarithmic multisine signal in the frequency domain for a given frequency band (in this example, $\left[ 0.01 \unit{rad/s}, 100 \unit{rad/s}\right]$ was used).
  In the previous section we saw that systems with equal damping and DC gain, will behave very similarly when excited by a log signal with spacing $\alpha$.

  Instead of using \eqref{eq:theoreticalBestVariationOmega} to establish the influence of the resonance frequency, we apply an equivalent approach that is more intuitive to interpret.

  \begin{figure}
    \centering
    \setlength\figureheight{5cm}
    \setlength\figurewidth{0.8\columnwidth}
    \input{\thisDir/figs/misalign.tikz}
    \caption[Depiction of the grid misalignment $\beta$ for a logarithmic generalized multisine.]{Graphical depiction of the misalignment coefficient $\beta$ with respect to a logarithmic frequency grid with spacing $\alpha$ for two systems with different damping.
    The $3\unit{dB}$ bandwidth (\sysBW) around the peak frequency $\wM$ of the system is highlighted.}
    \label{fig:misalignment-logms}
  \end{figure}

  We choose the resonance frequency of the systems such that
  \begin{equation}
   \wM\left( \wn \right)  = \omega_k + \beta \left( \omega_{k+1} - \omega_k \right)
   \text{ with } 0 \leq \beta < 1
   \label{eq:chosenVariationOmega}
  \end{equation}
  where $\omega_k$ and $\omega_{k+1}$ are the excited grid lines closest to the frequency $\wM$ where the FRF has its peak amplitude and $\beta$ is the misalignment coefficient that is varied over the interval $\RightOpenInterval{0,1}$ as depicted graphically in \figref{fig:misalignment-logms}
  This means that $\omega_k \leq \wM \leq \omega_{k+1}$ and that $\beta$ replaces the $\gamma$ factor in \eqref{eq:theoreticalBestVariationOmega}.
  If we take a look at the special case when $\beta=0$ (or $\beta=1$) in this expression, we see that this situation occurs when the peak value of the FRF at $\wM$ coincides with the grid line $\omega_k$ or $\omega_{k+1}$ respectively.
  Basically, we study second order systems whose peak frequency $\wM$ is shifted between two neighboring excited frequency lines.
  From the previous section, we know that the exact frequency matters little as long as it remains within the bulk of the frequency grid.
  In this particular example $\wn$ was chosen at approximately $1 \unit{rad/s}$.
